Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-06-06 paper/paper.tex
dust section and posterior

In this paper we use the noise-deconvolved \cmd\ as a prior to infer more precise parallax probability distribution functions (pdfs) from \gaia\ measurements.
Understanding the validity of this method can be challenging because we are using the data both for the prior and the likelihood in our inference.
To ease understanding, here we demonstrate the generaly methodology using a
simpler toy model with simulated data.
Instead of inferring parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior pdfs of the value $Y$.
We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel.
Similarly, noisy samples from this distribution are shown in the upper-right panel.
Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noise distribution.

@@

To infer a posterior pdf of the true $y$ value for each measured $y$ value we need a prior. We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y$ data themselves.
We learn the noise-deconvolved function using marginalized, maximum-likelihood estimation (MMLE), which marginalizes over the true data values and maximizes the likelihood of the data. Here our likelihood for each individual datum is
\begin{eqnarray}
p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \dd y_{\true, n} \\
&=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2) \quad,
\label{eq:toyLike}
\end{eqnarray}
where we have marginalized over the true sample $y_{\true, n}$. The resulting likelihood is a Gaussian because the noise model and our assumptions about the intrinsic width are both Gaussian. The likelihood of the full data set $\{y_n\}$ is then the product of the individual likelihoods
\begin{equation}
p(\{y_n\} \given m, b, t^2, \sigma_n^2) = \prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
\label{eq:toyLikeFull}
\end{equation}
Maximizing this joint likelihood of all the data is the same as minimizing
\begin{equation}
\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,\ln(2\pi\,[\sigma_n^2 + t^2])
\quad .
\label{eq:chisq}
\end{equation}

@@

In detail, the posterior points shown in the lower-right panel of
Figure~\ref{fig:toy} look more concentrated (vertically) than the true
values shown in the upper-left panel.
This is because the points in the lower-right (posterior) panel show
not typical values or typical draws from the posterior; they show the
modes (and, with error bars, the variances) of the posterior pdfs.
When an original data point has a large variance, the posterior ends
up close to the prior, which has a mode at the center line; the
posterior modes therefore concentrate more towards the center line
than any representative sampling.
A plot of posterior samplings (as opposed to a plot of posterior modes) would
show a scatter like that in the upper-left (truth) panel.

@@

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior pdf for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

Equation~\ref{eq:bayes} does not fully specify this procedure.
Because our prior comes from the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars. Specifically, we use the $J$ and $K$ band photometry from \tmass, as well as the parallax from \gaia. Our true, latent values are the $(J-K)^{\true}_n$ color and the $J$ band absolute magnitude $M^{\true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$. The observed data is the dust corrected color $(J-K)^{\corr}_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.

We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\data$ contains a transformation of the dust corrected absolute magnitude
\begin{equation}
y_0 = \varpi\,10^{\frac{1}{5}\,J^{\corr}_n} = 10^{\frac{1}{5}\,M^{\corr}_{J,n} + 2}
\label{eq:transform}
\end{equation}
where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain (Gaussian uncertainties are required to use \xd, see below).
Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
\begin{equation}
\begin{aligned}
\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K)^{\true}_n] \\
\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J- K)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
\mathrm{with \; apparent \; magnitude} \quad &J^{\corr}_n = J_n - Q_J\,E(B-V)_n \\
\mathrm{and \; color} \quad &(J - K)^{\corr}_n = J_n - K_n - Q_{JK}\,E(B-V)_n \\
\mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
\end{aligned}
\label{eq:data}
\end{equation}
where $E(B-V)_n$ is the output of the dust model of \cite{green15}, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.

We assume that the parallax likelihood is Gaussian, as well as the $(J-K)^{\corr}_n$ color likelihood. So our full likelihood for the $n$th star is
\begin{equation}
p(\data_n \given \truth_n, I_n) = \mathcal{N}(\data_n \given \truth_n, \covar_n)
\end{equation}
with \\
\[
\covar_n = \begin{bmatrix}
\sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^{\corr}}\right)^2
\end{bmatrix}
\]
where $\mathcal{N}(\mean, \mat{\Sigma})$ represents a normal distribution with
mean vector $\mean$ and covariance matrix $\mat{\Sigma}$.
Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude. We take $J_n^{\corr}$ as a point estimate and only propagate the parallax uncertainty.

@@

\section{Data and results}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{data.png}
\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K$ color, the $J$ band apparent magnitude, and the \gaia\ parallax. The blue lines represent the $1\sigma$ and $2\sigma$ contours for the distribution. To give a sense of the uncertainties, in the right panel we subsample the dataset and include the $1\sigma$ uncertainties. The uncertainties are dominated by the parallax noise, with many stars diverging to infinitely far away and therefore very intrinsically bright.}
\label{fig:data}
\end{figure}

@@

For the generation of the prior and the evaluation of the likelihood, we need to
correct the \tmass\ photometry for dust extinction.
With the emergence of 3D dust maps, it is now possible to estimate dust
corrections to stars within the Milky Way. The challenge is getting a
proper estimate of the distance to the star before inferring
$\varpi_{\true}$, since we need a dust correction prior to doing our
inference. Dust corrections are particularly non trivial for stars
with poor parallax signal to noise. Stars with poor signal to noise
have a likelihood that is consistent with infinite distance and can
therefore have severe dust corrections. This is most obvious in the
giant stars, which tend to have the lowest signal to noise, and can
therefore have dust corrections $> 2$ magnitudes. Although the dust
model is probabilistic, as well as our distance inference, to move
forward, we break from a fully Bayesian framework in which we might
infer the distance and the dust simultaneously (more about this in the
discussion). Instead, we correct for dust using a point estimate from
the 3D dust map \citep{green15} sampled at a distance estimated from
iteratively inferring and then sampling our posterior pdf over
$\varpi_{\true}$. In more detail, to apply a dust correction to the
\tmass\ photometry, we first generate our prior using the observed,
attenuated photometry. Using this raw prior, we infer more precise
parallaxes to all the stars in our sample. We then use this more
precise parallax posterior to get a measurement of dust in the 3D dust
map for each star. We then iterate this procedure. This approach minimizes severe
distances, and their associated severe dust corrections, which would
occur if using the likelihoods alone.

In detail, we take the $5\%$
quantile of each distance posterior (the closest part of the
distance posterior) and query the 3D dust map at that distance.
We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
correction because over-correcting for dust can push the peak of the likelihood
function into an unphysical, different space of the \cmd\ and thus bias the
parallax inference.
For each distance, we sample the median of the
probabilistic dust model. We apply each dust correction to our
\tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
= [0.709, 0.302]$ for bands $[J, K]$ respectively (see Table 6 in
\citealt{schlafly11}).
We then regenerate our prior using the dust-corrected photometry.
The dust values we obtain in our final
iteration, and which we apply to our photometry, are visualized in
galactic coordinates in \figurename~\ref{fig:dust}.
After applying the dust correction, we do not update the uncertainties, which is
technically wrong and under-represents the true uncertainties; as mentioned
previously, this leads to a less severe deconvolution.
In other words, the features of our inferred
  \cmd\ are conservatively broad (broader than if we over-estimated
  the errors which would have lead to excessive deconvolution and
  unrealistically narrow features). We could add the covariance of
the dust to our likelihoods, but this is beyond the scope of this
demonstration of concept. We iterate this process 10 times, however
the dust values seem to converge after about 2-3 iterations. The
iterative process allows for slight variations in distance and
corresponding dust to become more consistent with the \cmd.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{dust.png}
\caption{The converged dust values at the 5\% distance quantile in Galactic coordinates. This shows both the footprint of our analysis as well as the adopted extinction values. Each point represents a single star in our cross matched catalog. The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
\label{fig:dust}
\end{figure}

@@

Figure~\ref{fig:posterior} shows the end result of our inference for a few randomly-chosen objects: the posterior pdf on the true parallaxes.
We also show the likelihood and the prior.
In all cases, the prior information brought by the \cmd\ is significant, and leads to a compelling decrease in the parallax uncertainty.
The nature of our prior is also visible: it is apparent that it is made of a mixture of components, and that its impact strongly depend on the location of each object in \cmd\ space.

Figure~\ref{fig:posteriorCMD} shows the distribution of the posterior parallax inferences derived with our \cmd\ prior.
This is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes, and highlight the regions of \cmd\ space with the greatest improvement.
Those typically are the regions with the largest photometric and parallax errors, for the faintest objects.
This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posterior.png}
\caption{Five example posterior pdfs over parallax using our method.
The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
We chose example points that evenly spanned the color space, but within each color bin, we chose a point at random.
The remaining panels show the posterior pdf in blue, the prior pdf in green, and likelihood of the data in black.
Most stars have posterior pdfs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
Examples 1-4 are not very effected by the prior (green), where good data is good data.
Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior pdf, transformed into an absolute magnitude, remains on the main sequence.
Example 5 shows a significantly more precise posterior pdf compared with the likelihood, and looks very similar to the prior pdf.
It also exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
\label{fig:posterior}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posteriorCMD.png}
\caption{The distribution of expectation values of the parallax posterior pdfs transformed into an absolute magnitude vs color, the \cmd, in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution.
The right panel shows a subsample of data, plotting the expectation value and $1\sigma$ uncertainty of posterior pdf projected onto the \cmd.
Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd, shown in \figurename~\ref{fig:prior}
The visual of the prior is a sampling of the Gaussians, but this is the expectation
  value, so it lies much closer to the center of the
  distribution. Similar to the toy model, where the maximum of the
  posterior pdf of the noisy points lies closer to the center of the true distribution than the true values.}
\label{fig:posteriorCMD}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{delta.png}
\caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior relative to the likelihood as a function of the stellar color. The coloring of the points in both panels is by the fractional change in variance to guide the coloring of the right panel. The blue lines are the $1\sigma$ and $2\sigma$ contours. The median fractional change is 1.2. Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K)^{\corr}$ of about 0.6. The right panel shows the posterior point estimates on the CMD colored by fraction change in variance. There is significant changes for the red giant branch stars, as well as some main sequence stars.}
\label{fig:delta}
\end{figure}

@@

The method produces a mixture-of-Gaussian parallax posterior pdf for every
star in the intersection of \tgas, \tmass, and the dust map.
Associated with this paper is an electronic table containing our results, and everything
that is necessary to use them responsibly.
It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
photometry.
It contains posterior expectations and variances for the stellar parallax.
It also contains the full posterior pdf for each star, in the form of a set of pdf values,
evaluated on a common grid in parallax.
Because all of the data are public, and all of the code used for this project is available
publicly under an open-source license, power users can generate any of the figures in this
paper or any other data outputs straightforwardly.

\subsection{Distances to M67}
As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of 3--$5~{\rm Gyr}$.
Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
M67 is therefore a great test case for comparing parallax inferences with or without a prior.

>>>>

In this paper we use the noise-deconvolved \cmd\ as a prior to infer more precise parallax probability distribution functions (PDFs) from \gaia\ measurements.
Understanding the validity of this method can be challenging because we are using the data both for the prior and the likelihood in our inference.
To ease understanding, here we demonstrate the generaly methodology using a
simpler toy model with simulated data.
Instead of inferring parallax PDFs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer posterior PDFs of the value $Y$.
We perform this simpler inference using a prior that is also a 2D distribution: a 1D Gaussian in the y-direction, with a running mean $\mu = m\,x + b$ and some thickness $t$.
Here the thickness of the toy distribution mimics the true thickness or intrinsic scatter of the \cmd\ due to age, mass, and metallicity.
The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in \figurename~\ref{fig:toy}, upper-left panel.
Similarly, noisy samples from this distribution are shown in the upper-right panel.
Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief over $y_{\true,n}$, a true sample from the distribution, from $y_n$, a sample from the noise distribution.

@@

To infer a posterior PDF of the true $y$ value for each measured $y$ value we need a prior. We use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy $y$ data themselves.
We learn the noise-deconvolved function using marginalized, maximum-likelihood estimation (MMLE), which marginalizes over the true data values and maximizes the likelihood of the data. Here our likelihood for each individual datum is
\begin{eqnarray}
p(y_n \given m, b, t^2, \sigma_n^2) \ &=& \ \int \mathcal{N}(y_n \given y_{\true,n}, \sigma_n^2) \, \mathcal{N}(y_{\true,n} \given m\,x_n + b, t^2 ) \, \dd y_{\true, n} \\
&=&  \ \mathcal{N}(y_n \given m\,x_n + b, t^2 + \sigma_n^2) \quad,
\label{eq:toyLike}
\end{eqnarray}
where we have marginalized over the true sample $y_{\true, n}$. The resulting likelihood is a Gaussian because the noise model and our assumptions about the intrinsic width are both Gaussian. The likelihood of the full data set $\{y_n\}$ is then the product of the individual likelihoods
\begin{equation}
p(\{y_n\} \given m, b, t^2, \sigma_n^2) = \prod_n\, p(y_n \given m, b, t^2, \sigma_n^2) \quad .
\label{eq:toyLikeFull}
\end{equation}
Maximizing this joint likelihood of all the data is the same as minimizing
\begin{equation}
\chi^2 = \sum_n \frac{[y_n - [m\,x_n + b]]^2}{[\sigma_n^2 + t_n^2]} + \sum_n\,\ln(2\pi\,[\sigma_n^2 + t^2])
\quad .
\label{eq:chisq}
\end{equation}

@@

In detail, the posterior points shown in the lower-right panel of
Figure~\ref{fig:toy} look more concentrated (vertically) than the true
values shown in the upper-left panel.
This is because the points in the lower-right (posterior) panel show
not typical values or typical draws from the posterior; they show the
modes (and, with error bars, the variances) of the posterior PDFs.
When an original data point has a large variance, the posterior ends
up close to the prior, which has a mode at the center line; the
posterior modes therefore concentrate more towards the center line
than any representative sampling.
A plot of posterior samplings (as opposed to a plot of posterior modes) would
show a scatter like that in the upper-left (truth) panel.

@@

Under these assumptions, we would like to infer the true parallax, $\varpi_{\true}$, for each star given its measured parallax, $\varpi$, and associated uncertainty, $\sigma_{\varpi}$, from \tgas.
A first application of Bayes theorem leads to
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior PDF for the true
parallax, and $Z$ is a normalization constant (the model evidence, or marginalized likelihood).
The strength of our inference lies in how we generate our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This is because supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the \cmd.
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star $M$, when compared with its brightness or apparent magnitude $m$, contains distance information $d$ via the fundamental relation $m=M+5\log(d/10\, \mathrm{pc})$.
We include this in the inference by using the \cmd\ as part of our prior for $\varpi_{\true}$.
Specifically, we use the \cmd\ generated from the data themselves, as detailed in the next section.

Equation~\ref{eq:bayes} does not fully specify this procedure.
Because our prior comes from the \cmd, our inference will include not only the parallax from \gaia, but also the photometry of the stars. Specifically, we use the $J$ and $K_s$ band photometry from \tmass, as well as the parallax from \gaia. Our true, latent values are the $(J-K_s)^{\true}_n$ color and the $J$ band absolute magnitude $M^{\true}_{J,n}$ for each star, which we encapsulate in the true vector $\truth$. The observed data is the dust corrected color $(J-K_s)^{\corr}_n$, the \gaia\ parallax $\varpi_n$ in units of $[mas]$, and the dust corrected $J$ band magnitude $J^{\corr}_n$, which we encapsulate in the data vector $\data$.

We now describe a singularity of our method, which allows us to conservatively deconvolve the data into a \cmd\ prior without requiring a full hierarchical inference framework (which was explored in \citealt{leistedtHogg2017}).
Instead of transforming the parallax and apparent magnitude into an absolute magnitude, the data vector $\data$ contains a transformation of the dust corrected absolute magnitude
\begin{equation}
y_0 = \varpi\,10^{\frac{1}{5}\,J^{\corr}_n} = 10^{\frac{1}{5}\,M^{\corr}_{J,n} + 2}
\label{eq:transform}
\end{equation}
where, for this expression, it is important to remember that $\varpi$ is assumed to be in units of $[mas]$.
The transformation of the dust corrected absolute magnitude is required to to keep the uncertainties Gaussian, which, under some assumptions along with this transformation, we maintain (Gaussian uncertainties are required to use \xd, see below).
Our prior assumptions $I$ are the uncertainties in each of these measurements, as well as the 3D dust map. So our full terminology is
\begin{equation}
\begin{aligned}
\mathrm{true \, vector} \quad &\truth_n = [10^{0.2\,M^{\true}_{J,n} + 2}, (J-K_s)^{\true}_n] \\
\mathrm{data \, vector} \quad &\data_n = [\varpi_n\,10^{0.2J^{\corr}_n}, (J-K_s)^{\corr}_n] \;\hspace*{15mm}  n \in \mathrm{stars} \\
\mathrm{with \; apparent \; magnitude} \quad &J^{\corr}_n = J_n - Q_J\,E(B-V)_n \\
\mathrm{and \; color} \quad &(J-K_s)^{\corr}_n = J_n - K_{s,n} - Q_{JK}\,E(B-V)_n \\
\mathrm{under\ the\ assumptions} \quad &I_n = \left\{\left\{\sigma^2_{\varpi, n}\right\}, \left\{\sigma^2_{J,n}\right\}, \left\{\sigma^2_{K,n}\right\}, \mathrm{Dust \, Map}\right\}
\end{aligned}
\label{eq:data}
\end{equation}
where $E(B-V)_n$ is the output of the dust model of \cite{green15}, and $Q_{\lambda}$ is the correction factor for a photometric band assuming some reddening law.

We assume that the parallax likelihood is Gaussian, as well as the $(J-K_s)^{\corr}_n$ color likelihood. So our full likelihood for the $n$th star is
\begin{equation}
p(\data_n \given \truth_n, I_n) = \mathcal{N}(\data_n \given \truth_n, \covar_n)
\end{equation}
with \\
\[
\covar_n = \begin{bmatrix}
\sigma_{J,n}^2 + \sigma_{K,n}^2 & 0 \\
0 & \left(\sigma_{\varpi}\,10^{0.2\,J_n^{\corr}}\right)^2
\end{bmatrix}
\]
where $\mathcal{N}(\mean, \mat{\Sigma})$ represents a normal distribution with
mean vector $\mean$ and covariance matrix $\mat{\Sigma}$.
Here we have assumed that the parallax uncertainty is significantly larger than the $J$-band magnitude uncertainty and is therefore the dominant uncertainty in the transformed absolute magnitude. We take $J_n^{\corr}$ as a point estimate and only propagate the parallax uncertainty.

@@

\section{Data and results}
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{data.png}
\caption{The observed \cmd: To visualize the full dataset, in the left panel we use the point estimates of the \tmass\ $J-K_s$ color, the $J$ band apparent magnitude, and the \gaia\ parallax. The blue lines represent the $1\sigma$ and $2\sigma$ contours for the distribution. To give a sense of the uncertainties, in the right panel we subsample the dataset and include the $1\sigma$ uncertainties. The uncertainties are dominated by the parallax noise, with many stars diverging to infinitely far away and therefore very intrinsically bright.}
\label{fig:data}
\end{figure}

@@

For the generation of the prior and the evaluation of the likelihood, we need to
correct the \tmass\ photometry for dust extinction.
With the emergence of 3D dust maps, it is now possible to estimate dust
corrections to stars within the Milky Way. The challenge is getting a
proper estimate of the distance to the star before inferring
$\varpi_{\true}$, since we need a dust correction prior to doing our
inference. Dust corrections are particularly non trivial for stars
with poor parallax signal to noise; these stars have
a likelihood that is consistent with infinite distance and can
therefore have severe dust corrections. This is most obvious in the
giant stars, which tend to have the lowest signal to noise, and can
therefore have dust corrections $> 2$ magnitudes. Although the dust
model is probabilistic, as well as our distance inference, to move
forward, we break from a fully Bayesian framework in which we might
infer the distance and the dust simultaneously (more about this in the
discussion). Instead, we correct for dust using a point estimate from
the 3D dust map \citep{green15}. We determine this dust correction by
iteratively inferring, and then sampling, our posterior PDF over
$\varpi_{\true}$. We first generate our prior using the observed,
attenuated photometry. Using this raw prior, we infer more precise
parallaxes to all the stars in our sample. We use this more
precise parallax posterior to get a measurement of dust in the 3D dust
map for each star. We then iterate this process 10 times, however
the dust values seem to converge after about 2-3 iterations. The
iterative process allows for slight variations in distance and
corresponding dust to become more consistent with the \cmd.

In detail, we take the $5\%$
quantile of each distance posterior (the closest part of the
distance posterior) and query the 3D dust map at that distance.
We take the $5\%$ quantile---as apposed to the median---to do a minimal dust
correction because over-correcting for dust can push the likelihood
function into an unphysical, different space of the \cmd\ and thus bias the
parallax inference. At the 3D position of each star, we sample the \texttt{mode=median} value of the probabilistic dust map.
We apply each dust correction to our
\tmass\ photometry using Equation~\ref{eq:data}, where $Q_{\lambda}
= [0.709, 0.302]$ for bands $[J, K_s]$ respectively (see Table 6 in
\citealt{schlafly11}).
The dust values we obtain in our final
iteration, and which we apply to the photometry in our inference, are visualized in
galactic coordinates in \figurename~\ref{fig:dust}.
After applying the dust correction, we do not update the uncertainties, which is
technically wrong and under-represents the true uncertainties; as mentioned
previously, this leads to a less severe deconvolution.
In other words, the features of our inferred
  \cmd\ are conservatively broad (broader than if we over-estimated
  the errors which would have lead to excessive deconvolution and
  unrealistically narrow features). We could add the covariance of
the dust to our likelihoods, but this is beyond the scope of this
demonstration of concept.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{dust.png}
\caption{The converged dust values at the 5\% distance quantile in Galactic coordinates. This shows both the footprint of our analysis as well as the adopted extinction values. Each point represents a single star in our cross-matched catalog. The large missing areas are due to the \psone\ footprint ($>-30~{\rm deg}$ in declination)}
\label{fig:dust}
\end{figure}

@@

Figure~\ref{fig:posterior} shows the various probability distributions of our inference for a few randomly-chosen objects: the likelihood (black), prior (green), and posterior PDF (blue) on the true parallaxes.
For the majority of the \gaia\ stars, our prior slightly increases the precision of the posterior PDF compared with the \gaia\ variance. Our parallax posterior PDFs are more precise than the \tgas\ catalog entries by a median factor of $1.2$. This is visualized by Examples 1-4 in Figure~\ref{fig:posterior} which show slightly narrower posterior PDFs compared to the likelihoods.
However, $14\%$ of our parllax posterior PDFs are more precise than the \tgas\ catalog by a factor of $>2$. This larger change in precision is visualized in Example 5, which shows a significantly narrower posterior PDF compared to the likelihood.
The nature of our prior is also visible: it is apparent that it is made of a mixture of components, and that its impact strongly depend on the location of each object in \cmd\ space.

Figure~\ref{fig:posteriorCMD} shows the distribution of the posterior parallax PDFs derived with our \cmd\ prior.
This is further illustrated in Figure~\ref{fig:delta}, where we show the fractional changes, and highlight the regions of \cmd\ space with the greatest improvement.
Those typically are the regions with the largest photometric and parallax errors, for the faintest objects.
This is a fairly standard result: the deconvolution of uncertainties is stronger in noisier regions of the data, leading to narrower features in the upper part of the \cmd, for example.
\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posterior.png}
\caption{Five example posterior PDFs over parallax using our method.
The upper left panel shows the position of each example data point in black, with its observed uncertainties, plotted over our prior represented by a mixture of Gaussians in green.
We chose example points that evenly spanned the color space, but within each color bin, we chose a point at random.
The remaining panels show the posterior PDF in blue, the prior PDF in green, and likelihood of the data in black.
Most stars have posterior PDFs (blue) that are only slightly more precise than the data (black), which is well represented in this random sampling of points.
Examples 1-4 are not very effected by the prior (green), where good data is good data.
Example 4 is interesting because it lies in the same color range as the red clump, where there is a lot of support from the prior, but the likelihood does not have much support there so the parallax posterior PDF, transformed into an absolute magnitude, remains on the main sequence.
Example 5 shows a significantly more precise posterior PDF compared with the likelihood, and looks very similar to the prior PDF.
It also exemplifies that our prior tends to have the largest effect on stars in the red giant branch.}
\label{fig:posterior}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=\textwidth]{posteriorCMD.png}
\caption{The distribution of expectation values of the parallax posterior PDFs transformed into an absolute magnitude vs color, the \cmd, of the full data set is shown in the left panel, with the $1\sigma$ and $2\sigma$ contours of the distribution shown in black.
The right panel shows a subsample of data set, plotting the expectation value and $1\sigma$ uncertainty of posterior PDF projected onto the \cmd.
Notice that some parts of this \cmd\ are tighter than the prior, or noise-deconvolved \cmd, shown in \figurename~\ref{fig:prior}
The visual of the prior is a sampling of the Gaussians, but this is the expectation
  value, so it lies much closer to the center of the
  distribution. Similar to the toy model, where the maximum of the
  posterior PDF of the noisy points lies closer to the center of the true distribution than the true values.}
\label{fig:posteriorCMD}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{delta.png}
\caption{Fractional Change in Variance: The left panel shows the natural log of the fractional change in the variance of the posterior relative to the likelihood as a function of the stellar color. The coloring of the points in both panels is by the fractional change in variance to guide the coloring of the right panel. The blue lines are the $1\sigma$ and $2\sigma$ contours. The median fractional change is 1.2. Some posteriors have significantly smaller variances shown in black, but some actually increase in variance, shown as the yellow bump at $(J-K_s)^{\corr}$ of about 0.6. The right panel shows the posterior point estimates on the CMD colored by fraction change in variance. There is significant changes for the red giant branch stars, as well as some main sequence stars.}
\label{fig:delta}
\end{figure}

@@

The method produces a mixture-of-Gaussian parallax posterior PDF for every
star in the intersection of \tgas, \tmass, and the dust map.
Associated with this paper is an electronic table containing our results, and everything
that is necessary to use them responsibly.
It includes the \tgas\ identifier, the \gaia\ identifier, sky position, and dust-corrected
photometry.
It contains posterior expectations and variances for the stellar parallax.
It also contains the full posterior PDF for each star, in the form of a set of PDF values,
evaluated on a common grid in parallax.
Because all of the data are public, and all of the code used for this project is available
publicly under an open-source license, power users can generate any of the figures in this
paper or any other data outputs straightforwardly.

\subsection{Distances to M67}
As a test of the accuracy of these photometric parallax inferences, we inferred the parallaxes to stars in the open cluster M67.
M67 is at a Heliocentric distance of $800$--$900~{\rm pc}$ with an estimated age of $3$--$5~{\rm Gyr}$.
Previous work \citep{cbj15} has noted that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the parallax signal-to-noise (SN) is below 5, $\varpi / \sigma_\varpi < 5$.
With \gaia\ parallaxes having a minimum uncertainty of about $0.3~{\rm mas}$, the SN reaches 5 for the most certain parallax measurements at about $700~{\rm pc}$, just before the distance to this cluster.
M67 is therefore a great test case for comparing parallax inferences with or without a prior.
